---
title:               "BAG3 Pathway Analysis"
author:
  - name:            "Emir Turkes"
  - name:            "Columbia University"
date:                '`r strftime(Sys.time(), format = "%B %d, %Y")`'
bibliography:        "../bag3-pathway-analysis.bib"
biblio-style:        apalike
link-citations:      true
output:
  html_document:
    number_sections: true
    theme:           lumen
    toc:             true
    toc_depth:       2
    toc_float:
      collapsed:     false
      smooth_scroll: false

knit:
  (function(inputFile, encoding) {
    rmarkdown::render(
      inputFile,
      encoding = encoding,
      output_file = "../results/bag3-pathway-analysis.html"
    )
  })
---

```{r, include = FALSE}
# Copyright 2019 Emir Turkes
# 
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
# 
#     http://www.apache.org/licenses/LICENSE-2.0
# 
# Unless required by applicable law or agreed to in writing, software
# distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and
# limitations under the License.

knitr::opts_chunk$set(echo = TRUE)

library("magrittr")
```

<style type="text/css">

body{font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}

</style>

*This is an in-depth analysis based on one or more initial broad analyses.*

The background for this data is as follows:

- Co-IP MS (co-immunoprecipitation mass spectrometry) using BAG3 as bait from [github.com/eturkes/bag3-co-ip](https://github.com/eturkes/bag3-co-ip).

This analysis was performed in R except where noted.
The source code and instructions for rerunning the analysis can be found at [github.com/eturkes/bag3-pathway-analysis](https://github.com/eturkes/bag3-pathway-analysis).

# Final Results

**Read just this section for the final results of the analysis and a summary of the methods.**

# ~~~~ Breakdown of Methods ~~~~ {-}

The following sections breakdown the methods used to transform the contents in `Original Data` to those in `Final Results`.

# Original Data

This section contains the cleaned dataset generated from the analysis in [github.com/eturkes/bag3-co-ip](https://github.com/eturkes/bag3-co-ip).

```{r}
bag3_co_ip <- readRDS("../assets/bag3-co-ip.rds")
cols <- c("Genes", "Abundance Ratio: (scr/sh)", "scrRNA Abundance", "shBAG3 Abundance")

# From my own package, provides custom formatting.
eturkesRutils::datatable(bag3_co_ip) %>%
  DT::formatRound(2) %>% DT::formatStyle(cols[1], `text-align` = "center")
```

# Conclusion

This concludes the methods used in this analysis.
Below we simply write the data to disk for exploration in `Final Results` at the top of this report.

```{r}
devtools::session_info()
```

# References
