---
title:               "BAG3 Pathway Analysis"
author:
  - name:            "Emir Turkes"
  - name:            "Columbia University"
date:                '`r strftime(Sys.time(), format = "%B %d, %Y")`'
bibliography:        "../bag3-pathway-analysis.bib"
biblio-style:        apalike
link-citations:      true
output:
  html_document:
    number_sections: true
    theme:           lumen
    toc:             true
    toc_depth:       2
    toc_float:
      collapsed:     false
      smooth_scroll: false
knit:
  (function(inputFile, encoding) {
    rmarkdown::render(
      inputFile, encoding = encoding, output_file = "../results/bag3-pathway-analysis-report.html")})
---

```{r, include = FALSE}
# Copyright 2019 Emir Turkes
# 
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
# 
#     http://www.apache.org/licenses/LICENSE-2.0
# 
# Unless required by applicable law or agreed to in writing, software
# distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and
# limitations under the License.

knitr::opts_chunk$set(echo = TRUE, fig.width = 8.5)
```

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This is an in-depth analysis based on one or more initial broad analyses.*

The background for this data is as follows:

- Co-IP MS (co-immunoprecipitation mass spectrometry) using BAG3 as bait from [github.com/eturkes/bag3-co-ip](https://github.com/eturkes/bag3-co-ip).

This analysis was performed in R except where noted.
The source code and instructions for rerunning the analysis can be found at [github.com/eturkes/bag3-pathway-analysis](https://github.com/eturkes/bag3-pathway-analysis).

# Final Results

**Read just this section for the final results of the analysis and a summary of the methods.**

# ~~~~ Breakdown of Methods ~~~~ {-}

The following sections breakdown the methods used to transform the contents in `Original Data` to those in `Final Results`.
We start by loading in any required packages for our analysis and setting some global variables.

```{r}
library(conflicted)
library(DT)
library(gprofiler2)
library(devtools)
```

```{r}
assets_dir <- file.path(getwd(), "..", "assets")
results_dir <- file.path(getwd(), "..", "results")
```

# Original Data

This section contains the cleaned dataset generated from the analysis in [github.com/eturkes/bag3-co-ip](https://github.com/eturkes/bag3-co-ip).

```{r}
co_ip_genes <- readRDS(file.path(assets_dir, "co_ip_genes_final.rds"))
cols <- c("Genes", "Specificity: avg(Abun. Ratio, Watson)", "Abundance Ratio: (scr/sh)",
          "Watson Confidence", "scrRNA Abundance", "shBAG3 Abundance")
```

## Sort Schemes {.tabset}

Two sort schemes are presented, one to show specificity with BAG3 and another to show the most abundant Co-IP proteins.

### Specificity with BAG3

```{r}
co_ip_genes_spec <- co_ip_genes[order(-co_ip_genes[[cols[2]]], co_ip_genes[[cols[6]]]), ]
rownames(co_ip_genes_spec) <- NULL # Reset row numbers.
eturkesRutils::datatable(co_ip_genes_spec) %>% # From my own package, provides custom formatting.
  formatRound(c(2, 3, 4)) %>% formatStyle(cols[1], `text-align` = "center")
```

### Abundance of Protein

```{r}
co_ip_genes_abun <- co_ip_genes[order(-co_ip_genes[[cols[5]]], -co_ip_genes[[cols[6]]]), ]
rownames(co_ip_genes_abun) <- NULL # Reset row numbers.
eturkesRutils::datatable(co_ip_genes_abun) %>% # From my own package, provides custom formatting.
  formatRound(c(2, 3, 4)) %>% formatStyle(cols[1], `text-align` = "center")
```

## g:Profiler Analysis {.tabset}

We start by taking our gene lists and inputing them into [g:Profiler](https://biit.cs.ut.ee/gprofiler/gost), a web service for functional enrichment analysis [@raudvere_g:profiler:_2019].
The tool can be operated interactively through a web browser, but an R client is also provided, so we use that.
Since we have two gene lists, we shall continue analyzing them in parallel.
g:Profiler is the first step in a 2019 Nature Protocols paper on pathway analysis [@reimand_pathway_2019] that we will generally be following for this analysis.

### Specificity with BAG3

```{r}
gprofiler_spec <- gost(co_ip_genes_spec$Genes, organism = "rnorvegicus", ordered_query = TRUE)
gostplot(gprofiler_spec)
```

```{r, fig.width = 11, fig.height = 93}
publish_gosttable(gprofiler_spec)
```

### Abundance of Protein

```{r}
gprofiler_abun <- gost(co_ip_genes_abun$Genes, organism = "rnorvegicus", ordered_query = TRUE)
gostplot(gprofiler_abun)
```

```{r, fig.width = 11, fig.height = 94}
publish_gosttable(gprofiler_abun)
```

# Conclusion

This concludes the methods used in this analysis.
Below we simply write the data to disk for exploration in `Final Results` at the top of this report and output the `session_info`.

```{r}
session_info()
```

# References
